CrystalPlan: an Experiment Planning Tool for Crystallography

### 2010-12-16 docs/CrystalPlan_paper/paper.tex
Refs #182. All Pete's comments are in.

To this end, CrystalPlan was written in Python \cite{python}, a powerful, open,
cross-platform scripting language. Python was chosen for its ease of cross-platform deployment, and for
its attractive language features. The numpy and scipy libraries (two
open-source, commonly installed Python libraries) were used for many of the
calculations \cite{numpy}. Some of the critical calculations were also written
with inline C code, which speed up some calculations by up to $400\times$ as compared with
the pure Python version (which can still be used, in case of incompatibilities).

@@

The calculation code of CrystalPlan is placed in a model layer and kept
completely separate from eventual GUI elements. This makes it possible to run
calculations using Python scripts, if desired. Unit tests for each part of the
calculations are used to ensure consistent results on different platforms and
systems.

@@

Peaks were integrated using ISAW, and the resulting .integrate file was loaded
into CrystalPlan for comparison. The predicted and actual measurement
positions on each detector face were compared for the 554 integrated peaks
found. Very good agreement was found; the
RMS difference between the predicted and actual measurement wavelength was only
0.016 \ang. Inspecting the detector position offsets showed that the main
contributor to this error was a slight misalignment of the goniometer, which
CrystalPlan can help diagnose.

@@

We have described CrystalPlan, a software program designed to quickly produce
efficient experiment plans for wide bandwidth time-of-flight Laue
diffractometers. This program will help maximize scientific productivity at SNS
beamlines such as TOPAZ.

>>>>

To this end, CrystalPlan was written in Python \cite{python}, a powerful, open,
cross-platform scripting language. Python was chosen for its ease of cross-platform deployment, and for
its attractive language features. The numpy and scipy libraries (two
open-source, commonly installed Python libraries) were used for many of the
calculations \cite{numpy,scipy}. Some of the critical calculations were
also written with inline C code, which speed up some calculations by up to $400\times$ as compared with
the pure Python version (which can still be used, in case of incompatibilities).

@@

The calculation code of CrystalPlan is placed in a model layer and kept
completely separate from eventual GUI elements, using a Model-View-Controller
architecture \cite{Gamma95}. This makes it possible to run calculations using
Python scripts, if desired. Unit tests for each part of the calculations are used to ensure
consistent results on different platforms and  systems.

@@

Peaks were integrated using ISAW, and the resulting .integrate file was loaded
into CrystalPlan for comparison. The predicted and actual measurement
positions on each detector face were compared for the 554 integrated peaks
found. Very good agreement was found; the
RMS difference between the predicted and actual measurement wavelength was only
0.016 \ang. The offsets in the peak position were plotted, which
helped show that the main contributor to this error was a slight misalignment of
the goniometer.

@@

We have described CrystalPlan, a software program designed to quickly produce
efficient experiment plans for wide bandwidth time-of-flight Laue
diffractometers. This program will help increase scientific productivity at SNS
beamlines such as TOPAZ.
